---
title: "Understanding the Concept of Mean in Statistics"
output: 
  learnr::tutorial:
    css: ["custom.css"]
    includes:
      in_header: "header.html"
    theme: cosmo
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: "A comprehensive interactive tutorial on understanding and interpreting the concept of mean in statistics"
---

```{r setup, include=FALSE}
library(learnr)
library(ggplot2)
library(dplyr)
library(plotly)
library(knitr)
library(DT)
knitr::opts_chunk$set(
  echo = FALSE,
  fig.width = 8,
  fig.height = 5,
  fig.align = 'center',
  warning = FALSE,
  message = FALSE,
  comment = ""
)

# Custom styling for better visuals
tutorial_options(
  exercise.timelimit = 60,
  exercise.checker = NULL
)
```

## Welcome to the Mean Tutorial {.tabset .tabset-fade .tabset-pills}

### Overview

Welcome to this **comprehensive tutorial** on understanding the concept of Mean in statistics!

**What you'll master:**

- **Definition of Mean**

- **Types of Mean**

- **Applications of Mean**

- **Visualizing Mean**

- **Interpreting Mean in Context**

### Learning Objectives

By the end of this tutorial, you will be able to:

- Define and explain the concept of mean

- Differentiate between types of mean (arithmetic, geometric, harmonic)

- Calculate mean for different datasets

- Visualize mean using plots

- Interpret mean in real-world scenarios

## Understanding the Mean

### What is Mean?

The **mean** is a measure of central tendency that represents the average value of a dataset. It is calculated by summing all the values and dividing by the total number of values.

**Formula:**

$$
\mu = \frac{\sum x_i}{n}
$$

where:

- $\mu$ is the mean

- $x_i$ are the individual data points

- $n$ is the total number of data points

### Real-World Example

Consider the following dataset of test scores: **85, 90, 78, 92, 88**.

- Sum of scores: $85 + 90 + 78 + 92 + 88 = 433$

- Total number of scores: $n = 5$

- Mean: $\mu = \frac{433}{5} = 86.6$

The mean test score is **86.6**.

### Visualizing the Mean

```{r mean_plot, fig.cap="Visualization of test scores and their mean. The red line represents the mean value."}
# Test scores
test_scores <- c(85, 90, 78, 92, 88)

# Create data frame
scores_data <- data.frame(
  Score = test_scores,
  Student = factor(1:5)
)

# Create plot
ggplot(scores_data, aes(x = Student, y = Score)) +
  geom_bar(stat = "identity", fill = "#3498db", alpha = 0.7) +
  geom_hline(yintercept = mean(test_scores), color = "#e74c3c", linetype = "dashed", size = 1.2) +
  labs(title = "Test Scores and Mean",
       subtitle = "The red dashed line represents the mean value",
       x = "Student", y = "Score") +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.line = element_line(color = "black")
  )
```

### Types of Mean

There are different types of mean used in statistics:

- **Arithmetic Mean**: The simple average of a dataset.

- **Geometric Mean**: The nth root of the product of n values.

- **Harmonic Mean**: The reciprocal of the arithmetic mean of reciprocals.

### Example Table

| Type of Mean      | Formula                          | Use Case                          |
|-------------------|----------------------------------|------------------------------------|
| Arithmetic Mean   | $\mu = \frac{\sum x_i}{n}$      | General average                   |
| Geometric Mean    | $\mu_g = \sqrt[n]{\prod x_i}$   | Growth rates                      |
| Harmonic Mean     | $\mu_h = \frac{n}{\sum \frac{1}{x_i}}$ | Rates and ratios                  |

**Explanation:**

This table summarizes the different types of mean, their formulas, and common use cases.

### Quiz Time!

```{r mean_quiz}
quiz(
  question("What is the formula for the arithmetic mean?",
    answer("$$\\mu = \\frac{\\sum x_i}{n}$$", correct = TRUE,
           message = "Correct! The arithmetic mean is calculated by summing all values and dividing by the total number of values."),
    answer("$$\\mu = \\sqrt[n]{\\prod x_i}$$"),
    answer("$$\\mu = \\frac{n}{\\sum \\frac{1}{x_i}}$$"),
    answer("None of the above"),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),

  question("Which type of mean is best for calculating growth rates?",
    answer("Arithmetic Mean"),
    answer("Geometric Mean", correct = TRUE,
           message = "Correct! The geometric mean is commonly used for growth rates."),
    answer("Harmonic Mean"),
    answer("None of the above"),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```

### Practice Exercise

```{r mean_practice, exercise=TRUE}
# Practice calculating the mean
# Dataset: 12, 15, 20, 25, 30

# 1. Calculate the sum of the dataset
# 2. Count the total number of values
# 3. Divide the sum by the total number to find the mean
```

```{r mean_practice-solution}
# Solution
# Dataset: 12, 15, 20, 25, 30

# Sum: 12 + 15 + 20 + 25 + 30 = 102
# Total number: 5
# Mean: 102 / 5 = 20.4

cat("The mean is 20.4")
```
